$K$-corrections and filter transformations \\ in the ultraviolet, optical, and near infrared

### 2006-08-01 docs/paper/kcorrect.tex
paper changes
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@560 75dff457-8751-0410-ab1a-cd701654ce6b

There are limitations to all stellar population synthesis libraries,
and our results are therefore restricted to be at best as good as our
input models. There are some well-known areas of uncertainty in the
stellar populations models in the UV and IR. In the UV, old
populations sometimes contain extreme horizontal branch stars that
have lost their envelopes and are thus extremely hot, that dominate
the flux at $\lambda \sim 1500$ \AA. There is certainly no physical
model that predicts when this happens, and no consensus about how
these stars should be accounted for in isochrone synthesis (that is,
what fraction of stars undergo this phase as a function of mass and
metallicity). In the near-IR, the spectra of thermally-pulsating
asymptotic giant branch stars may dominate the flux in certain
intermediate age populations (2--3 Gyr). Again, the fraction of stars
in this phase and what their spectra are is under dispute. {\bf what
does BC03 do?}  We do not address these uncertainties here, though we
note that for low redshift galaxies our results from BC03 do explain
most galaxies' colors (though see below for our discussion of the high
redshift GOODS sample). {\bf refs}

@@

The templates do well on higher redshift data, as well. For example,
Figure \ref{goods} shows the color residuals for GOODS data. Here we
find that there are somewhat larger residuals with respect to the
uncertainties (again shown as the thick dashed lines). We can compare
this result to that in Figure \ref{goods_special}, which we get by
fitting a new set of templates to {\it only} the GOODS data (this
alternate set of five templates is also distributed in the manner that
Appendix \ref{format} describes). From Figure \ref{goods_special}, we
conclude that {\it some} of the errors are irreducible, and require
either a larger set of templates for high redshift galaxies (a
possibility we find a priori unlikely), or that there are simply
errors in the catalogs or the input filter curves. For example, in the
$z$ and $V$ bands we find significant trends between redshifts $z=0.5$
and $1.5$ (echoed in Figure \ref{goods}). More significantly, in the
$H$ band there is at least a 10\% offset in the magnitudes across all
redshifts, which is almost certainly a catalog or filter curve
error. Other errors (such as a large scatter in the residuals in the
$B$ band and some redshift dependence in the $J$ band) are clearly due
to the fact that the templates are not primarily designed for GOODS
data. All in all, we recommend using the special templates for
$K$-corrections within the GOODS data set.

@@

Consider Figures \ref{twomass_resid} and \ref{twomass_predicted}. The
former shows the color residuals of {\it fitting} to both SDSS and
2MASS data for each galaxy. The latter shows the color residuals when
we fit {\it only} to the SDSS data and do not include 2MASS in the fit
at all. Without any input from 2MASS the templates do a very good job
of predicting the 2MASS fluxes, with a scatter of 20--30\% --- not far
from the uncertainties in the 2MASS fluxes themselves. Of course, it
is not surprising that it is easy to predict the 2MASS fluxes ---
stellar spectra are very simple in the NIR, so having the $ugriz$-band
fluxes from SDSS yields much information regarding the redder
bandpasses.

However, we will note here that this result flies in the face of a
persistent insistence that NIR observations are necessary to measure
stellar masses of galaxies. If we can predict the NIR observations
themselves, clearly they cannot be adding much to our knowledge of the
underlying stellar mass. This fact changes one's decisions about what
data is best to use for calculating the stellar mass function.  Basing
it on 2MASS data only improves slightly the stellar mass estimate of
each galaxy, while restricting (and thus biasing) the sample
significantly at the lowest luminosities and surface brightnesses
relative to the SDSS. Of course, if at high redshift galaxies have
a very different locus of their SEDs, then perhaps this statement is
less true for them.

@@

In addition to $K$-corrections, these template fits also provide
physical interpretations of the galaxies, since they correspond to
actual stellar population synthesis models. Indeed, the code outputs
three parameters relevant to the star-formation history: the current
mass of stars in the galaxy; the stellar mass-to-light ratio of the
galaxy; and the fraction of the total amount of star-formation that
has occurred recently. We do not believe that fitting just to
broad-band colors can ever give extremely accurate values for these
quantities.  However, in this section

@@

Finally, we can also calculate measures of the recent star-formation
rate from these fits. Since our fits are to broad-band photometry, and
not based on emission line measurements, we cannot expect to have
detailed information about the very recent star-formation
rate. However, we can try to measure what fraction of the total
star-formation has occurred in, say, the past 1 Gyr:
\begin{equation}
b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
\end{equation}
where the definition here is meant to be similar to the birthrate
parameter $b$ in \citet{kennicutt94a} and references therein.  Note
that here $b_G$ is defined in terms of the total integrated
star-formation history (unlike the stellar masses referred to above)
as is standard for the $b$ parameter. Figure \ref{umr_bg} shows the
relationship between the rest-frame $u-r$ color for SDSS galaxies and
$b_G$. For galaxies with $u-r < 2.5$ (redder than which galaxies are
mostly reddened, edge-on disks) these two parameters follow a simple
relationship:
\begin{equation}
\log_{10} b_g = -0.55 (u-r). 
\end{equation}
Note, however, the large scatter (about 0.3 dex at 1$\sigma$).

We do not want to overstate the validity of these physical
interpretations. They are sufficiently good physical interpretations
to explain broad band data, but are not uniquely so. Certainly more
detailed spectroscopic analysis can provide better constraints on
star-formation histories and stellar masses.

>>>>

There are limitations to all stellar population synthesis libraries,
and our results are therefore restricted to be at best as good as our
input models. There are some well-known areas of uncertainty in the
stellar populations models in the UV and IR. In the UV, old
populations sometimes contain extreme horizontal branch stars that
have lost their envelopes and are thus extremely hot, that dominate
the flux at $\lambda \sim 1500$ \AA. There is certainly no physical
model that predicts when this happens, and no consensus about how
these stars should be accounted for in isochrone synthesis (that is,
what fraction of stars undergo this phase as a function of mass and
metallicity). While the models of \cite{bruzual03a} do contain
post-AGB and planetary nebula phases that provide FUV flux in old
populations, they do not contain extreme horizontal branch stars or
other candidates that might provide more of the ``UV-excess'' seen in
elliptical galaxies (\citealt{greggio90a, brown00a}).  In the near-IR,
the spectra of thermally-pulsating asymptotic giant branch (TP-AGB)
stars may dominate the flux in certain intermediate age populations
(2--3 Gyr). With respect to these phases, \citet{bruzual03a} use
models that are calibrated to observations of TP-AGB stars
nearby. \citet{maraston05a} show that a different choice of TB-AGB
models can produce significantly higher flux in the NIR at
intermediate ages. Thus, the choices for AGB and post-AGB evolution we
make here by using the models \cite{bruzual03a} are not definitive,
but we do not address these controversies here and they remain an
uncertainty in the physical interpretation of galaxy fluxes.  We do
note that for low redshift galaxies our results from BC03 do explain
most galaxies' colors (though see below for our discussion of the high
redshift GOODS sample).

@@

The templates do well on higher redshift data, as well. For example,
Figure \ref{goods} shows the color residuals for GOODS data. Here we
find that there are somewhat larger residuals with respect to the
uncertainties (again shown as the thick dashed lines). We can compare
this result to that in Figure \ref{goods_special}, which we get by
fitting a new set of templates to {\it only} the GOODS data (this
alternate set of five templates is also distributed in the manner that
Appendix \ref{format} describes). From Figure \ref{goods_special}, we
conclude that {\it some} of the errors are irreducible, and require
either a different set of templates for high redshift galaxies or that
there are simply errors in the catalogs or the input filter
curves. For example, in the $z$ and $V$ bands we find significant
trends between redshifts $z=0.5$ and $1.5$ (echoed in Figure
\ref{goods}). More significantly, in the $H$ band there is at least a
10\% offset in the magnitudes across all redshifts, which is almost
certainly a catalog or filter curve error. One possibility for these
problems in the red are that intermediate age populations have flux in
the near-IR from TP-AGB that our model omits; for example,
\citet{maraston05a} has shown that such stars can significantly change
the near-IR colors. We defer a fuller investigation of these
discrepancies.  Other errors (such as a large scatter in the residuals
in the $B$ band and some redshift dependence in the $J$ band) are
clearly due to the fact that the templates are not primarily designed
for GOODS data. All in all, we recommend using the special templates
for $K$-corrections within the GOODS data set.

@@

Consider Figures \ref{twomass_resid} and \ref{twomass_predicted}. The
former shows the color residuals of {\it fitting} to both SDSS and
2MASS data for each galaxy. The latter shows the color residuals when
we fit {\it only} to the SDSS data and do not include 2MASS in the fit
at all. Without any input from 2MASS the templates do a very good job
of predicting the 2MASS fluxes, with a scatter of 20--30\% --- not far
from the uncertainties in the 2MASS fluxes themselves.

However, we will note here that this result flies in the face of the
view of many that NIR observations are necessary to measure stellar
masses of galaxies. If we can predict the NIR observations themselves,
clearly they cannot be adding much to our knowledge of the underlying
stellar mass. This fact changes one's decisions about what data is
best to use for calculating the stellar mass function.  Basing it on
2MASS data only improves slightly the stellar mass estimate of each
galaxy, while restricting (and thus biasing) the sample significantly
at the lowest luminosities and surface brightnesses relative to the
SDSS. Of course, if at high redshift galaxies have a very different
locus of their SEDs (for example, as perhaps is true of the GOODS
galaxies in the NIR), then perhaps this statement is less true for
them.

@@

In addition to $K$-corrections, these template fits also provide
physical interpretations of the galaxies, since they correspond to
actual stellar population synthesis models. Indeed, the code outputs
three parameters relevant to the star-formation history: the current
mass of stars in the galaxy; the stellar mass-to-light ratio of the
galaxy; and the fraction of the total amount of star-formation that
has occurred recently. We do not believe that fitting just to
broad-band colors can ever give extremely accurate values for these
quantities. In this section we demonstrate consistency between our
results and those of other groups for the same galaxies.

@@

Finally, we can also calculate measures of the recent star-formation
rate from these fits. Since our fits are to broad-band photometry, and
not based on emission line measurements, we cannot expect to have
detailed information about the very recent star-formation
rate. However, we can try to measure what fraction of the total
star-formation has occurred in, say, the past 1 Gyr:
\begin{equation}
b_{G} = \frac{\int_0^{1\mathrm{~Gyr}} dt\, \mathrm{SFR}(t)}
{\int_0^{a_{\mathrm{Univ}}} dt\, \mathrm{SFR}(t)}
\end{equation}
where the definition here is meant to be similar to the birthrate
parameter $b$ in \citet{kennicutt94a} and references therein.  Note
that here $b_G$ is defined in terms of the total integrated
star-formation history (unlike the stellar masses referred to above)
as is standard for the $b$ parameter. Figure \ref{umr_bg} shows the
relationship between the rest-frame $u-r$ color for SDSS galaxies and
$b_G$. For galaxies with $u-r < 2.5$ (redder than which galaxies are
mostly reddened, edge-on disks) these two parameters follow a simple
relationship:
\begin{equation}
\log_{10} b_g = -0.55 (u-r). 
\end{equation}
Note, however, the large scatter (about 0.3 dex at 1$\sigma$).  The
red dots show the results of using the formula of \citet{hopkins03a}
to infer the current star-formation rate from the $u$-band,
integrating over 1 Gyr, scaling by the integrated stellar mass from
the {\tt kcorrect} determination, and then multiplying by four (which
is the minimum appropriate correction for dust extinction in $u$
according to \citealt{hopkins03a}). Our results roughly scale as those
of \citet{hopkins03a} do.

We do not want to overstate the validity of these physical
interpretations. They are sufficiently good physical interpretations
to explain broad band data, but are not uniquely so. Certainly more
detailed spectroscopic analysis can provide better constraints on
star-formation histories and stellar masses.  However, the results
from {\tt kcorrect} are roughly in agreement with what other authors
find with comparable data.
